Asymmetric tail-dependence in extreme climatic events affects population-extinction risk

### 2019-08-30 MainText.Rmd
abstract edited

With global climate change, extreme climatic events (ECEs) appear to be more common than usual. These ECEs 
could pose severe impact on the resistance or resilience of a population or community level through
life-history traits of the species. Therefore predicting the impact of ECEs 
on extinction risk of a population is necessery. We here introduced the potential use of "tail-dependence" 
property of a copula in modeling the extinction risk of an age-structured single patch model. We also 
extended the generalized findings of density-independent metapopulation model from @Ghosh_copula 
to density-dependence metapopulation models with this copula approach. For metapopulation models, 
we showed extinction risk not only depended on the nature of tail-dependence of spatially correlated 
noise but was also influenced by the inherent dynamics of the population-model.

\textbf{\textit{Keywords:}} extinction risk, extreme events, tail dependence, copula, density-dependent model

>>>>

With global climate change, extreme climatic events (ECEs) become more frequent. These ECEs 
could pose severe impact on the resistance or resilience of a population and/ or community level through
life-history traits of the species. Therefore predicting the impact of extreme events 
on population extinction risk is necessery. We here introduce the potential use of "tail-dependence" 
property of two correlated ecological variables in modeling the extinction risk of 
an age-structured single patch model. We also extend this idea 
to density-dependent metapopulation models. For metapopulation models, 
we show extinction risk not only depends on the nature of tail-dependence of 
spatially correlated noise but is also influenced by the inherent dynamics 
of the population-model. To illaborate, we demonstrate left tail dependent noise across spatially correlated patches 
cause higher extinction risk than the right tail dependent noise, when the model show undercompensatory dynamics.
But in overcompensatory regime, the trend is opposite. Our study paves the path for future research considering this tail-dependence instead of simple correlation in population viability analysis.

\textbf{\textit{Keywords:}} extinction risk, extreme events, tail dependence, density-dependent model
